### 2017-06-17 Doc/classicthesis/Chapters/Chapter0A.tex
UML sequence diagram
First draft
segundo borrador

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}
\chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

% TODO
In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed. The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system. The second one is the issuance of the credential, divided in the three REST calls needed in the delegation. The last one is the proving for the Presentation Policy from the Verifier, using two REST calls.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{gfx/APDUsInitIoTSC}
	\end{center}
	\caption{Init IoT Smart Card APDU Commands exchanged.}
	\label{fig:APDUsInitIoTSC}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.78\linewidth]{gfx/IssuanceAPDUs}
	\end{center}
	\caption{Issuance APDU Commands.}
	\label{fig:IssuanceAPDUs}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/APDUsProving}
	\end{center}
	\caption{Proving APDU Commands.}
	\label{fig:APDUsProving}
\end{figure}

\chapter{Guide on how to continue the project development}

>>>>

%********************************************************************
% Appendix
%*******************************************************
% If problems with the headers: get headings in appendix etc. right
%\markboth{\spacedlowsmallcaps{Appendix}}{\spacedlowsmallcaps{Appendix}}

\chapter{Test: APDU Commands exchanged}\label{ch:resultsdiagrams}

In this appendix we provide 3 sequence diagrams highlighting the APDU Commands exchanged during our testbed.

The first one is the setup of the IoT smart card, storing the system parameters needed to work in the deployed P2ABCE system. The first Command is called \texttt{isAndroid} because the \texttt{HardwareSmartcard} class distinguishes Android phones acting as smart cards from MULTOS smart cards, in order to use Extended APDUs. Then configures the smart card, setting the PIN (``\texttt{1234}'' by default) and PUK. Finally, it copies the cryptographic system parameters to the smart card with multiple \texttt{SET} instructions.

The second image shows the issuance of the credential, divided in the three REST calls needed during the delegation. If we recall Idemix's issuance protocol, the User performed, during his first step, an exponentiation with his secret key and a ZKP, which can be identified by the \texttt{COMMITMENT} and \texttt{ISSUANCE RESPONSE} commands. During the second step, the User only has to verify the Issuer's ZKP and store the credential if the signature is valid. Because the P2ABC Engine can perform Verification of ZKP without the need of secret keys stored in the smart card, the APDU Commands in this phase are only for storing the credential information inside the IoT smart card's BLOB\footnote{(\textbf{B}inary \textbf{L}arge \textbf{Ob}jects} database, with the \texttt{STORE BLOB} instruction.

The last diagram represents the proving for the Presentation Policy received from a Verifier, using two REST calls. During the first REST call, the Service stores data for the proving in the smart card, but before starting the proving itself, the device must choose an identity for the proving. This is done with the Identity Service, that in the PoC chooses the first identity available. After this, the second REST call starts the proving, creating the \textit{commitment} and \textit{responses} that integrate any ZKP, and depends on the device's secret keys\footnote{We, again, refer the reader who wants to know more about ZKPs to \citep{tfgmates}.}.

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.9\linewidth]{gfx/UML/APDUsInitIoTSC}
	\end{center}
	\caption{Init IoT Smart Card APDU Commands exchanged.}
	\label{fig:APDUsInitIoTSC}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=0.78\linewidth]{gfx/UML/IssuanceAPDUs}
	\end{center}
	\caption{Issuance APDU Commands.}
	\label{fig:IssuanceAPDUs}
\end{figure}

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/APDUsProving}
	\end{center}
	\caption{Proving APDU Commands.}
	\label{fig:APDUsProving}
\end{figure}

\chapter{Attached code structure}\label{ch:code}

Given the size of the project, we identify four main directory structures to arrange the code.

First, the Dockerfiles, scripts that automatize the creation of Docker images. These images are like virtual machine snapshots, although Docker is not a virtualization system, and we use the images to create containers, that are like new virtual machines started from the snapshots. We use two Dockerfiles, one for the P2ABCE compilation environment, with Java and Maven, and another one for building the Omega2 binaries, using the LEDE SDK.

\dirtree{%
	.1 Docker/. 
	.2 P2ABCE/. 
	.3 Dockerfile. 
	.3 idemix-3.0.36-binaries/. 
	.2 LEDE SDK/. 
	.3 Dockerfile. 
}

\hfil

The original P2ABCE project is of considerable size, but we list here the most interesting directories that we had to modify to make the project compatible with IoT smart cards.

\hfil

\dirtree{%
	.1 P2ABCE/. 
	.2 Code/. 
	.3 core-abce/. 
	.4 abce-components/. 
	.5 src/main/java/eu/abc4trust/smartcard/. 
	.6 IoTsmartcardio/.
	.7 util/. 
	.8 IoTsmartcardConnection.java.  
	.7 IoTCard.java. 
	.7 IoTCardChannel.java. 
	.7 IoTCardTerminal.java. 
	.6 HardwareSmartcard.java. 
	.6 SoftwareSmartcard.java. 
	.4 abce-services/. 
	.5 src/main/java/eu/abc4trust/services/. 
	.6 UserService.java. 
	.6 VerificationService.java. 
	.6 IssuanceService.java. 
}

\hfil

The IoT smart card toolkit is the core of this project. Managed with CMake, we see the \texttt{CMakeLists.txt} files, including the \texttt{Toolchain-omega2-mipsel.cmake} file with the cross-compiler information. We divide the project between the \textit{source} C files, and the header C files. Inside these directories, the organization is as stated in the implementation chapter: \textit{core} or \textit{common}, \textit{util interfaces} and \textit{external utilities}.

\hfil

\dirtree{%
	.1 p2abce\_iot\_toolkit/. 
	.2 CMakeLists.txt. 
	.2 Toolchain-omega2-mipsel.cmake. 
	.2 util\_sc\_tests/. 
	.2 util\_sc/. 
	.3 CMakeLists.txt. 
	.3 BIOSC.c. 
	.3 p2abc\_iot\_toolkit\_src/. 
	.4 smartcard\_common/. 
	.5 APDU\_handler.c. 
	.5 global\_vars.c. 
	.5 m\_adapted\_API.c. 
	.5 subroutines.c. 
	.4 smartcard\_external\_utilities/. 
	.5 cJSON.c. 
	.4 smartcard\_utils\_interface/. 
	.5 APDU\_IO\_util.c. 
	.5 arithmetic\_util.c. 
	.5 crypto\_util.c. 
	.5 serialize\_util.c. 
	.5 system\_funcs.c. 
	.3 p2abc\_iot\_toolkit\_include/. 
	.4 error\_codes.h. 
	.4 macrologger.h. 
	.4 smartcard\_common/. 
	.5 APDU\_handler.h. 
	.5 APDU\_types.h. 
	.5 abc4T\_types.h. 
	.5 defs\_consts.h. 
	.5 defs\_errs.h. 
	.5 defs\_ins.h. 
	.5 defs\_types.h. 
	.5 global\_vars.h. 
	.5 m\_adapted\_API.h. 
	.5 subroutines.h. 
	.4 smartcard\_external\_utilities/. 
	.5 base64.h. 
	.5 cJSON.h. 
	.4 smartcard\_utils\_interface/. 
	.5 APDU\_IO\_util.h. 
	.5 arithmetic\_util.h. 
	.5 crypto\_util.h. 
	.5 serialize\_util.h. 
	.5 system\_funcs.h. 
}

\hfil

Finally, the test scripts, including a Python file to test the BIOSC protocol, a directory with the testbed scripts (one file to orchestrate every device from one terminal), and a more didactic version with a script per device.

\hfil

\dirtree{%
	.1 scripts/. 
	.2 testBIOSC.py. 
	.2 tests/\DTcomment{One test script}. 
	.3 setupTest.sh. 
	.3 testNoIoT.sh. 
	.3 IoTtest.sh. 
	.2 distributed-PoC/\DTcomment{Distributed test}. 
	.3 Omega2/. 
	.3 RaspberryPi3/. 
	.3 ThirdPartyServer/. 
}

### 2017-06-17 Doc/classicthesis/Chapters/Conclusions.tex
segundo borrador

Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work.

@@

@@

	\item Integrate the IoT+Idemix solution in other system

	Our seventh objective, that we couldn't get to design in depth. The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system. The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).

@@

@@

% Último objetivo: FIWARE

>>>>

Recalling the objectives listed in \autoref{objectives:section}, we think we accomplish them, except the implementation of a PoC in the most constrained device possible, where we used LEDE to ease our development, and the last objective, where we could have given a use to the PoC, and we leave as part of the future work. In spite of this negative auto-critic, we have to acknowledge that this is the first privacy-preserving ABC implementation for IoT, designed for extensibility, interoperability and maintenance.

Finally, we are very thankful to IBM's grant for the \textit{Privacy Preserving Identity Management applied to IoT} project, which allowed us to get this far. Also, this paper will be presented for publication in a JCR magazine.

@@

	\item Use cryptographic hardware

	Atmel offers many solutions for secure memory and cryptographic operations chips, with serial interfaces for the IoT devices. We could benefit from this chips by storing the secret keys of the credentials inside the secure memory, and speed up the SHA256 and AES128 calculations, currently software implemented and slow compared to any hardware implementation.

@@

	\item Integrate the IoT+Idemix solution in FIWARE

	Our sixth objective, that we couldn't get to design in depth. The original idea was to integrate in a FIWARE\footnote{FIWARE \url{https://www.fiware.org/}} environment the Fiware Keyrock and Idemix Issuer, so both, users and IoT devices' identities are managed in the IdM\footnote{Identity Management - KeyRock \url{https://catalogue.fiware.org/enablers/identity-management-keyrock}} system. The Issuer generates credentials based on the user’s attributes held in the Keyrock, and IoT devices can authenticate against Fi-ware services (using Idemix), or authenticate against other IoT devices (M2M).

@@

	\item Research P2ABCE policies for IoT applications

	The next step, after integrating Idemix with an Identity Management system, should be to design more complex applications to IoT scenarios. We talked about a smart building in the introduction, but the straightforward proofs we can present depend on the signed values in the credentials. The Zero-Knowledge Proofs cryptography is already implemented, we only have to extend the protocols to benefit from them.

	For example, a combination of authentication and arbitrary proof could be achieved by a combined proof of knowledge of a signed credential (current authentication system) and said arbitrary proof, although the values used in it aren't actually signed by the Issuer, the first proof gives credibility.

	Another solution for the same problem, similar to a PKI scheme, the IoT device receives a signed credential including an attribute which is his own public key. The IoT device can issue new credentials with the parameters it needs for the arbitrary proofs. The Verifier then certifies the new credential is signed by the IoT device's public key.

@@

### 2017-06-17 Doc/classicthesis/Chapters/Design.tex
First draft
segundo borrador

In this chapter we will describe the process for defining how a constrained IoT device may be integrated in a system like P2ABCE. We will also describe the PoC implementation carried out to test a realistic deployment.

@@

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
	\end{center}
	\caption{IoT in P2ABCE Architecture.}
	\label{fig:P2ABCE-IoT}
\end{figure}

@@

	In \autoref{fig:P2ABCE-IoT} the IoT device is represented with two interfaces, physical or virtual. One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc. Through this interface, the P2ABCE XML messages are exchanged as in any other deployment. This allows an IoT device to interact with other actors without special adaptations to the protocol. The other interface allows a secure communication with the delegation server. Both the delegation messages and the APDU Dialogue are transmitted over this interface, making it a point of attack to the system, and we will talk about its security in the delegation process.

@@

	The IoT device starts an interaction with another actor, e.g. an Issuer or Verifier, receiving a P2ABCE XML file.

@@

	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys.

@@

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/DelegationProving}
	\end{center}
	\caption{IoT Delegation in P2ABCE for Proving.}
	\label{fig:DelegationProving}
\end{figure}

@@

In this PoC the only P2ABCE Service that needed to be modified was the User Service. We added the following REST call

\begin{center}
	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
\end{center}

where we communicate the P2ABCE server that an IoT Smart Card is accessible via \textit{host} and \textit{port}. Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.

@@

The implementation of this simple protocol is done in both Java and C, one for the P2ABCE server and the other one for the IoT smart card.

@@

\textbf{TODO} :  diagrama secuencia de la reimplementación de javax.smartcardio y su uso por HardwareSmartcard

@@

The \texttt{main.h} file, from the original ABC4Trust project, also implemented some functions that were equivalent to the ones available in \texttt{multos.h}. We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified. This last detail is very important, because in \citep{vullers2013efficient} they noted that MULTOS' \texttt{ModularExponentiation} function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform. After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}.

@@

For example, Atmel's ATAES132A \citep{ATAES132A}
\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
offers a serial chip for secure key storage, AES128 execution and random number generation. Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption. For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.

>>>>

In this chapter we describe the design of our proposal to integrate IoT constrained devices as part of the privacy preserving system P2ABCE. The ultimate goal is to enable constrained IoT devices to play the Idemix User role, interacting autonomously in order to authenticate and demonstrate their credential attributes in a privacy-preserving fashion.

@@

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/P2ABCE-IoT-bw}
	\end{center}
	\caption{Proposed high level Architecture for integrating IoT devices in P2ABCE.}
	\label{fig:P2ABCE-IoT}
\end{figure}

@@

	In \autoref{fig:P2ABCE-IoT} shows our proposed architecture, in which the IoT device is represented with two interfaces, physical or virtual. One allows external communications to other machines, including other P2ABCE actors, that could be on the Internet, a corporate LAN, a M2M overlay network, etc. Through this interface, the P2ABCE XML messages are exchanged as in any other deployment. This allows an IoT device to interact with other actors without special adaptations to the protocol. The other interface allows a secure communication with the delegation server. Both the delegation messages and the APDU Dialogue are transmitted over this interface, making it a point of attack to the system, and we will talk about its security in the delegation process.

@@

	The IoT device, acting as a User, starts an interaction with another actor, e.g., against an Issuer to obtain a signed credential, or against a Verifier to demonstrate certain property of its attributes in a privacy-preserving way.

@@

	The server may need to send APDU Commands to the IoT smart card to read the credential information or perform cryptographic operations involving private keys, necessarily stored inside the IoT device.

	These APDU Commands include a list of 70 different instructions, identified by the APDU INS byte. Some of them manage the smart card, like changing the PIN code, and many are P2ABCE specific, like storing the system cryptographic parameters. Some examples:

	\begin{tabular}{|c|l|}
		\hline
		INS & Description \\
		\hline
		\texttt{0x52} & \texttt{SET\_CREDENTIAL} \\ & Sets a new credential inside the smart card.\\
		\texttt{0x5A} & \texttt{GET\_CREDENTIAL\_PUBLIC\_KEY} \\ & Returns a stored credential's public key.\\
		\texttt{0x60} & \texttt{GET\_PRESENTATION\_COMMITMENT} \\ & First cryptographic operation during Proving.\\
		\texttt{0x62} & \texttt{GET\_PRESENTATION\_RESPONSE} \\ & Second cryptographic operation during Proving.\\
		\hline		
	\end{tabular}

@@

\begin{figure}[bth]
	\begin{center}
		\includegraphics[width=\linewidth]{gfx/UML/DelegationProving}
	\end{center}
	\caption{Designed interactions exchanged during the IoT delegation for P2ABCE Proving operation.}
	\label{fig:DelegationProving}
\end{figure}

@@

In this PoC, only the P2ABCE's User Service needed to be modified, because this is the role the IoT device plays. We added the following REST call:
\begin{center}
	\textit{/initIoTsmartcard/{issuerParametersUid}?host=\&port=}
\end{center}
where we communicate the P2ABCE server that an IoT Smart Card is accessible via the IP address \textit{host} and TCP port \textit{port}. Then a new \textit{HardwareSmartcard} object is stored in the P2ABC Engine, but instead of the \textit{javax.smartcardio} Oracle's \textit{CardTerminal} implementation, we use our own \textit{IoTsmartcardio} implementation for the \textit{HardwareSmartcard} constructor, that we will discuss in the following section.

@@

The implementation of this simple protocol is done in Java for the P2abce delegation server and in C for the IoT smart card device.

@@

@@

The \texttt{main.h} file, from the original ABC4Trust project, also implemented some functions that were equivalent to the ones available in \texttt{multos.h}. We replaced them for the standard functions in the C-API, checking that the MEL code did the same as the documentation specified. This last detail is very important, because in \citep{vullers2013efficient} they noted that MULTOS' \texttt{ModularExponentiation} function does not accept exponents larger than the modulus size, so they implemented \texttt{SpecialModularExponentiation}, dividing the exponentiation in two that MULTOS could perform. After this analysis, when those particularities appeared, our reimplementation of the MULTOS API accepts the \textit{expanded functionality}. In total, we reduced to eighteen the number of MULTOS functions to reimplement, and with future refactoring this number may decrease.

@@

For example, Atmel's ATAES132A\footnote{ATAES132A 32K AES Serial EEPROM  Datasheet - \url{http://ww1.microchip.com/downloads/en/DeviceDoc/Atmel-8914-CryptoAuth-ATAES132A-Datasheet.pdf}}
\marginpar{\includegraphics[width=0.9\linewidth]{gfx/atmel} \\ Atmel's cryptography chips.}
offers a serial chip for secure key storage, AES128 execution and random number generation. Another serial chip like ESP8266 offers WiFi connectivity, typically used with Arduino, and can also perform AES encryption. For random number generation, a technique used with Contiki devices is to read from sensors aleatory data and use it as seed.

### 2017-06-17 Doc/classicthesis/Chapters/Introduction.tex
segundo borrador

The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, breafly, we can think of it as a set of devices, mainly resource constrained, that are interconnected between them in order to achieve a goal. For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.

@@

With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination. Thats why current solutions rely on  \acp{ZKP}
\marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}.}
, cryptographic methods that allow to proof knowledge of some data without disclosing it.

@@

The goal of this project is to integrate Idemix with the IoT. It will be done using the ABC4Trust's \ac{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}. This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

>>>>

The \ac{IoT} is a term with a wide range of interpretations \citep{Atzori20102787}, breafly, we can think of it as billions of devices, mainly resource constrained, that are interconnected between them, and the Internet, in order to achieve a goal. For example, a network of lampposts with proximity sensors that talk to each other so they light up part of the street when a passerby walks by, but save energy when not; greenhouses with automated irrigation to balance costs and quality vegetables; or cars exchanging traffic data to reduce city pollution from traffic jams or avoid accidents.

@@

With usual symmetric and asymmetric cryptography it seems impossible to create such credentials, without an explosion of signatures over every possible combination. For this reason, current solutions rely on  \acp{ZKP}
\marginpar{To read more about ZKPs, refer to my Mathematics thesis \citep{tfgmates}}
, cryptographic methods that allow to proof knowledge of some data without disclosing it.

@@

%%%
So far, Idemix or privacy-ABCs have been successfully applied to deal with traditional Internet scenarios, in which users can authenticate and prove their attributes against Service provider. However, due to the reduced computational capabilities of certain IoT devices, it has not been yet considered for IoT scenarios. As we study in the state of the art chapter, current implementations are based on Java, which requires high computational and memory resources to be executed, and to the best of our knowledge, this is the first proposal that tries to apply an IoT solution for privacy-preserving authentication and authorization, based on Anonymous credential Systems, like Idemix.
%%5

The goal of this project is to integrate Idemix with the IoT. It will be done using the ABC4Trust's \acf{P2ABCE}, a framework that defines a common architecture, policy language and data artifacts for an attribute based ecosystem, cryptographically based on either IBM's Idemix or Microsoft's U-Prove \footnote{P2ABCEngine \url{https://github.com/p2abcengine/p2abcengine}}. This gives us a standardized language to exchange Idemix's messages between IoT devices and any other P2ABCE actor.

Once the IoT devices can execute Idemix, the new step is to take advantage of it in other deployments. In a smart building, different privacy policies could protect sensitive data, like how many people there are, and their identities, where a thermostat would only need to know if the amount of people is between some boundaries, but in case of emergency, the police department could request all the information available. We think that the first approach to achieve these ideas is first to integrate Idemix in known IoT identity systems, like FIWARE project.
